Type-Safe Modular Parsing

### 2016-11-24 paper/sections/Overview.tex
Text

trait ExprParser {
  lexical.delimiters += ("(", ")")
  val pVar: PackratParser[Expr] =
    ident ^^ Var
  val pApp: Fix[PackratParser[Expr]] =
    e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
  val pVarApp: Fix[PackratParser[Expr]] =
    e => pApp(e) | pVar | "(" ~> e <~ ")"
}
\end{lstlisting}

@@

val p = new ExprParser {}.pVarApp
parseAndPrint(p)("x y")
\end{lstlisting}

@@

\begin{lstlisting}
trait LambdaParser extends ExprParser {
  lexical.delimiters += ("\\", ".")
  val pLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
    e => "\\" ~> ident ~ ("." ~> e) ^^ { case x ~ b => Lam(x, b) }
  val pVarAppLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
    e => pLam(e) | pVarApp(e)
}
\end{lstlisting}

@@

\begin{lstlisting}
val p = new LambdaParser {}.pVarAppLam
parseAndPrint(p)("(\\x.x) y")
\end{lstlisting}

@@

\begin{lstlisting}
trait VarApp[E] {
  def Var(x: String): E
  def App(e1: E, e2: E): E
}
\end{lstlisting}

>>>>

trait ExprParser {
  lexical.delimiters += ("(", ")")

  val pVar: PackratParser[Expr] =
    ident ^^ Var
  val pApp: Fix[PackratParser[Expr]] =
    e => e ~ e ^^ { case e1 ~ e2 => App(e1, e2) }
  val pVarApp: Fix[PackratParser[Expr]] =
    e => pApp(e) | pVar | "(" ~> e <~ ")"
}
\end{lstlisting}

@@

parseAndPrint(new ExprParser {}.pVarApp)("x y")
\end{lstlisting}

@@

\begin{lstlisting}
trait LambdaParser extends ExprParser {
  lexical.delimiters += ("\\", ".")

  val pLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
    e => "\\" ~> ident ~ ("." ~> e) ^^ { case x ~ b => Lam(x, b) }
  val pVarAppLam: (=> PackratParser[Expr]) => PackratParser[Expr] =
    e => pLam(e) | pVarApp(e)
}
\end{lstlisting}

@@

\begin{lstlisting}
parseAndPrint(new LambdaParser {}.pVarAppLam)("(\\x.x) y")
\end{lstlisting}

@@

\begin{lstlisting}
trait Alg[E] {
  def Var(x: String): E
  def App(e1: E, e2: E): E
}
\end{lstlisting}

Then we are able to define operations over the syntax in an extensible way. Here we write a pretty printing operation for this language.

\begin{lstlisting}
trait Print extends Alg[String] {
  def Var(x: String) = x
  def App(e1: String, e2: String) = "(" + e1 + " " + e2 + ")"
}
\end{lstlisting}

And also parser for it as below. Notice the parsing function \lstinline{pVarApp} takes an argument of type \lstinline{Alg[E]}, that means it accepts any instance of the algebra interface \lstinline{Alg}.

\begin{lstlisting}
trait Parser[E] {
  lexical.delimiters += ("(", ")")

  val pVarApp: Alg[E] => Fix[PackratParser[E]] = alg => e =>
    e ~ e ^^ { case e1 ~ e2 => alg.App(e1, e2) } | ident ^^ alg.Var | "(" ~> e <~ ")"
}
\end{lstlisting}

We glue them together by defining an object.

\begin{lstlisting}
object VarApp {
  trait Alg[E] {..}
  trait Print extends Alg[String] {..}
  trait Parser[E] {..}
}
\end{lstlisting}

To use the parser, we must provide an algebra instance as the operation to construct the parsing result. As shown in the code below, we feed the pretty printing operation \lstinline{VarApp.Print} to the generic \lstinline{parse} function. It will print \lstinline{(x y)} as the result.

\begin{lstlisting}
def parse[E](inp: String)(alg: VarApp.Alg[E]): E = {
  val p = new VarApp.Parser[E] {}.pVarApp(alg)
  val t = phrase(fix(p))(new lexical.Scanner(inp))
  if (t.successful) t.get else scala.sys.error(t.toString)
}

def parseAndPrint(inp: String) = println(parse(inp)(new VarApp.Print{}))

parserAndPrint("x y")
\end{lstlisting}
